Relative parabolic factorizations of Chevalley groups

### 2015-03-22 dennis-vaserstein.tex

Let $i$, $j$ be two distinct indices $1\leq i,j \leq n$. From Levi decomposition it follows that the following four subsets of $\E(\Phi, R, I)$ are equal to each other:
\begin{multline}\nonumber \U(\Phi, I)\cdot \U(\Phi^-, I) \cdot \E(\Delta_i, R, I) \cdot \EP_j(R, I) = \U(\Sigma_i, I)\cdot \U(\Sigma^-_i, I) \cdot \E(\Delta_i, R, I) \cdot \EP_j(R, I) = \\
= \EP_i(R, I) \cdot \E(\Delta_j, R, I) \cdot \U(\Sigma_j^-, I)\cdot \U(\Sigma_j, I) = \EP_i(R, I) \cdot \U(\Sigma^-_i\cap \Sigma^-_j, I) \cdot \EP_j(R, I). \end{multline}
Denote by $A_{ij}$ any of the equal subsets from the previous formula. 
Clearly, $A_{ij}$ is invariant with respect to conjugation by root subgroups $X_{\alpha\ssub{k}}$ for $1\leq k\leq n$ and $X_{-\alpha\ssub{k}}$ for $k\neq i,j$.

@@

\begin{lemma}\label{Stein_reduction} 
Suppose that there exists a subset $\widetilde{L_r} \subseteq \E(\Delta_r, R, I)$ satisfying the following two conditions:
\begin{enumerate}[label=(\alph*)] 
 \item\label{stein_cond1} One has $\U(\Phi^+, I)\cdot \U(\Phi^-, I) \cdot \widetilde{L_r} = \U(\Phi^+, I)\cdot \U(\Phi^-, I) \cdot \E(\Delta_r, R, I).$
 \item\label{stein_cond2} There is an inclusion $X_{-\alpha\ssub{r}}(I)^{\widetilde{L_r}} \subseteq \EP_s(R, I).$
\end{enumerate}
Then the set $A_{rs}$ is invariant with respect to
\begin{itemize}
 \item conjugation by elements of $\EP_s$;
 \item left multiplication by elements of $X_{-\alpha\ssub{r}}(I)$.
\end{itemize} \end{lemma}
\begin{proof} Set $A':=\U(\Phi^+, I)\cdot \U(\Phi^-, I) \cdot \widetilde{L_r} \cdot \EP_s(\Phi, R, I).$
From condition~\ref{stein_cond1} of the lemma it follows that $A_{rs}=A'$.

Let us check the first statement of the lemma. The subgroup $\EP_s$ is generated by root subgroups $X_{\alpha\ssub{i}}$ for $1\leq i\leq n$ and $X_{-\alpha\ssub{i}}$ for $i\neq s$,
therefore we only need to show that $A'$ is invariant with respect to conjugation by $X_{-\alpha\ssub{r}}$.

From Lemma~\ref{lemma:dv_unipotent} it follows that
\begin{multline}\nonumber A'^{X_{-\alpha\ssub{r}}} \subseteq (\U(\Sigma_r\setminus \{\alpha_r\}, I) \cdot \U(\Phi^-, I) \cdot X_{\alpha\ssub{r}}(I) \cdot X_{-\alpha\ssub{r}}(I) \cdot \widetilde{L_r} \cdot \EP_s(R, I))^{X_{-\alpha\ssub{r}}} \subseteq \\
\subseteq \U(\Phi^+, I) \cdot \U(\Phi^-, I) \cdot (X_{\alpha\ssub{r}}(I) \cdot X_{-\alpha\ssub{r}}(I) \cdot \widetilde{L_r})^{X_{-\alpha\ssub{r}}} \cdot \EP_s(R, I). \end{multline}
On the other hand, from condition~\ref{stein_cond2} it follows that:
\begin{multline}\nonumber 
X_{\alpha\ssub{r}}(I) \cdot X_{-\alpha\ssub{r}}(I) \cdot \widetilde{L_r} \subseteq
X_{\alpha\ssub{r}}(I) \cdot \widetilde{L_r} \cdot \EP_s(R, I)) \subseteq \\ \subseteq
\widetilde{L_r} \cdot \U(\Sigma_r^+, I) \cdot \EP_s(R, I) \subseteq \widetilde{L_r} \cdot \EP_s(R, I). \end{multline}
Thus we obtain the first statement of the lemma:
\begin{multline}\nonumber 
 A'^{X_{-\alpha\ssub{r}}} \subseteq \U(\Phi^+, I) \cdot \U(\Phi^-, I) \cdot \widetilde{L_r} ^{X_{-\alpha\ssub{r}}} \cdot \EP_s(R, I) \subseteq \\ \subseteq
 \U(\Phi^+, I) \cdot \U(\Phi^-, I) \cdot \widetilde{L_r} \cdot \EP_s(R, I) \cdot \EP_s(R, I) = A'. \end{multline}

The proof of the second part of the lemma is similar to the case just analyzed. \end{proof}

>>>>

\subsection{Main reduction.}
Let $i$, $j$ be two distinct indices $1\leq i,j \leq n$. From Levi decomposition it follows that the following four subsets of $\E(\Phi, R, I)$ are equal to each other:
\begin{multline}\nonumber \U(\Phi, I)\cdot \U(\Phi^-, I) \cdot \E(\Delta_i, R, I) \cdot \EP_j(R, I) = \U(\Sigma_i, I)\cdot \U(\Sigma^-_i, I) \cdot \E(\Delta_i, R, I) \cdot \EP_j(R, I) = \\
= \EP_i(R, I) \cdot \E(\Delta_j, R, I) \cdot \U(\Sigma_j^-, I)\cdot \U(\Sigma_j, I) = \EP_i(R, I) \cdot \U(\Sigma^-_i\cap \Sigma^-_j, I) \cdot \EP_j(R, I). \end{multline}
Denote by $A_{ij}$ any of the equal subsets from the previous formula. 
Clearly, $A_{ij}$ is invariant with respect to conjugation by root subgroups $X_{\alpha\ssub{k}}$ for $1\leq k\leq n$ and $X_{-\alpha\ssub{k}}$ for $k\neq i,j$.

@@

\begin{lemma}\label{lemma:conj_lemma} Let $G$ be a group, $H\leq G$ be a subgroup and $A, B\subseteq G$ be two subsets.
Assume that $A$ is invariant with respect to left multiplication by elements of $B$ and conjugation by elements of the subgroup $H$.
Then $A$ is invariant with respect to left multiplication by elements of the set $B^H = \{b^h \vert b\in B, h\in H\}$. \end{lemma}
\begin{proof} $b^h \cdot A \subseteq h^{-1}\cdot b \cdot h \cdot A \subseteq h^{-1} \cdot b \cdot A \cdot h \subseteq h^{-1} \cdot A \cdot h \subseteq A. \qedhere$ \end{proof}

\begin{lemma}\label{Stein_reduction} 
Suppose that there exists a subset $\widetilde{L_r} \subseteq \E(\Delta_r, R, I)$ satisfying the following two conditions:
\begin{enumerate}[label=(\alph*)] 
 \item\label{stein_cond1} One has $\U(\Phi^+, I)\cdot \U(\Phi^-, I) \cdot \widetilde{L_r} = \U(\Phi^+, I)\cdot \U(\Phi^-, I) \cdot \E(\Delta_r, R, I).$
 \item\label{stein_cond2} There is an inclusion $X_{-\alpha\ssub{r}}(I)^{\widetilde{L_r}} \subseteq \EP_s(R, I).$
\end{enumerate}
Then $A_{rs} = \E(\Phi, R, I)$. \end{lemma}
\begin{proof} Set $A:=\U(\Phi^+, I)\cdot \U(\Phi^-, I) \cdot \widetilde{L_r} \cdot \EP_s(\Phi, R, I).$
From condition~\ref{stein_cond1} of the lemma it follows that $A_{rs}=A$.

The first step of the proof is to show that $A_{rs}$ is invariant with respect to conjugation by the root subgroup $X_{-\alpha\ssub{r}}$ 
and left multiplication by elements of $X_{-\alpha\ssub{r}}(I)$. To prove this, observe first that from Lemma~\ref{lemma:dv_unipotent} and condition~\ref{stein_cond2} it follows that
\begin{multline}\nonumber 
A \subseteq \U(\Sigma_r\setminus \{\alpha_r\}, I) \cdot \U(\Phi^-, I) \cdot X_{\alpha\ssub{r}}(I) \cdot X_{-\alpha\ssub{r}}(I) \cdot \widetilde{L_r} \cdot \EP_s(R, I) \subseteq \\ 
\subseteq \U(\Sigma_r\setminus\{\alpha_r\}, I) \cdot \U(\Phi^-, I) \cdot X_{\alpha\ssub{r}}(I) \cdot X_{-\alpha\ssub{r}}(I) \cdot \widetilde{L_r} \cdot \EP_s(R, I) \subseteq \\
\subseteq \U(\Sigma_r\setminus\{\alpha_r\}, I) \cdot \U(\Phi^-, I) \cdot \widetilde{L_r} \cdot \U(\Sigma_r, I) \cdot \EP_s(R, I)  \cdot \EP_s(R, I) \subseteq \\
\subseteq \U(\Sigma_r\setminus\{\alpha_r\}, I) \cdot \U(\Phi^-, I) \cdot \widetilde{L_r} \cdot \EP_s(R, I). \end{multline}
Now it remains to apply Lemma~\ref{lemma:dv_unipotent} once again:
\begin{equation}\nonumber A^{X_{-\alpha\ssub{r}}} \subseteq \U(\Phi^+, I) \cdot \U(\Phi^-, I) \cdot \widetilde{L_r} ^{X_{-\alpha\ssub{r}}} \cdot \EP_s(R, I) \subseteq A. \end{equation}
\begin{equation}\nonumber X_{-\alpha\ssub{r}}(I) \cdot A \subseteq \U(\Phi^+, I) \cdot X_{-\alpha\ssub{r}}(I) \cdot \U(\Phi^-, I) \cdot \widetilde{L_r} \cdot \EP_s(R, I) \subseteq A. \end{equation}

  Notice that the subgroup $\EP_s$ is generated by root subgroups $X_{\alpha\ssub{i}}$ for $1\leq i\leq n$ and $X_{-\alpha\ssub{i}}$ for $i\neq s$,
therefore in fact we have just proved that $A$ is invariant with respect to conjugation by all elements of $\EP_s$.

Now we are ready to prove the needed equality. By virtue of Theorem~\ref{theorem:Stepanov} it suffices to show that
$A$ is invariant with respect to left multiplication by elements of the following two families:
\begin{itemize}
 \item $z_{\alpha}(s, \xi)$, $s\in I$, $\xi \in R$, $\alpha\in\Sigma^-_s$;
 \item $x_{\beta}(s)$, $s \in I$, $\beta \in \Phi$. \end{itemize}
In fact, thanks to Lemma~\ref{lemma:conj_lemma} it suffices to prove this statement only for the second family.
Moreover, we can additionally assume $\beta\in \Sigma_r^-$.
%TODO: To be continued 
\end{proof}
